
```{r,include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE}

require(RODBC)
require(rgdal)
require(devtools)
require(roxygen2)
require(geosphere)
require(SpatialHub)
require(lubridate)
require(bio.utilities)
require(bio.lobster)
require(rstanarm)
require(rlang)
require(glue)
require(PBSmapping)

 
	p = bio.lobster::load.environment()
	la()
	assessment.year = 2023 ##Check Year
	p$syr = 1989
  p$yrs = p$syr:assessment.year



figdir = file.path("LFA35_Update")
    
	    p$lfas = c("35") # specify lfa
    	p$subareas = c("35") # specify subareas for data summary
	   

## from Oracle
#lobster.db('logs.redo') 
#lobster.db('temperature.data.redo')
#lobster.db('fsrs.redo')
groundfish.db('odbc.redo',datayrs = 1970:2020)

	
#Not Oracle		
logs=lobster.db('process.logs.redo')
logs=lobster.db("process.logs")
land = lobster.db('annual.landings')
land = lobster.db('seasonal.landings')
land=lobster.db('seasonal.landings.redo')

```
# Analysis and Response

#Primary Indicator
The LFA 35-38 assessment (DFO 2020) provided a full analysis of stock health by describing fishery performance and indicators for abundance and biomass, fishing pressure and reproduction. From this assessment commercial Catch Per Unit Effort (CPUE) has been identified as the primary indicator.

##Catch Per Unit Effort 
Catch rates are a preferred indicator over landings data as they are standardized to account for the level of fishing effort, which is especially important in effort controlled fisheries. However, they vary during the fishing season due to changes in biomass and catchability, which can be accounted for in catch rate models.

Data for assessing catch rates primarily comes from mandatory logs that were not put in place until the mid 2000’s. This time series covers the current high productivity period and a lower productivity period from 2006-2010. The median of the high productivity period (2011-2018 was used as the proxy for the biomass at carrying capacity (K). Following the recommendations of DFO (2009), the USR and LRP were set to 40% and 20% of the K proxy. The 3-year running median is used to compare the commercial catch rates to the USR and LRP.  This value will dampen the impact of any anomalous years, which may occur due to factors outside of changes in abundance.

```{r CPUE1, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE, fig.cap="Time series of commercial catch rates (black line) for LFA 35, along with the 3-year running median (blue line). The horizontal lines represent the upper stock (dashed green line) and limit reference point (dotted red line)."}

		logs=lobster.db("process.logs")
		TempModelling = TempModel( annual.by.area=F, redo.data=F)
			CPUE.data<-CPUEModelData(p,redo=T,TempModelling)

			## Commercial CPUE MOdels
			mf1 = formula(logWEIGHT ~ fYEAR + DOS + TEMP + DOS * TEMP)

			CPUE.data<- CPUEModelData(p,redo=F)
			t=with(subset(CPUE.data,DOS==1),tapply(TEMP,LFA,mean))
		
			pData=list()

			CPUEModelResults = list()

			for(i in 1:length( p$lfas)){

				mdata = subset(CPUE.data,LFA==p$lfas[i]&SYEAR%in%p$yrs)
				if(nrow(mdata)>10){
				CPUEModelResults[[i]] = CPUEmodel(mf1,mdata,t=t[i],d=1)
				pData[[i]]=CPUEModelResults[[i]]$pData
				pData[[i]]$LFA=p$lfas[i]
			}
			}

			names(CPUEModelResults) = p$lfas

			CPUEindex=do.call("rbind",pData)


	# plot

	tiff("C:/Users/Howsevj/Documents/LFA35_Update/plotcpue.tiff", width=7,height=6, units="in",res=300)

	l35 = subset(CPUEindex,LFA==35,c("YEAR","mu"))
	k = median(l35$mu[which(l35$YEAR %in% 2011:2018)])
	usr = .4*k
	lrp = .2*k
	

plot(l35$YEAR,l35$mu,xlab='Year',ylab='CPUE (kg/TH)',type='p',pch=16,ylim=c(0,6))
	running.median = with(rmed(l35$YEAR,l35$mu),data.frame(YEAR=yr,running.median=x))
	l35=merge(l35,running.median,all=T)
	lines(l35$YEAR,l35$running.median,col='blue',lty=1,lwd=3)
	abline(h=usr,col='green',lwd=2,lty=2)
	abline(h=lrp,col='red',lwd=2,lty=3)
	
dev.off()
```

The CPUE trend indicates an increase in stock biomass occured between 2005 and 2011. The CPUE time series has remained high (more than twice the USR)



#Secondary Indicators


#Landings and Effort



```{r landings and effort, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE, fig.cap="Time series of landings (bars), and effort (sold line with points) for LFA 35. The data for 2018-2019 fishing season is incomplete (COLOUR)."}
	
	land = lobster.db('seasonal.landings')
	land$YR = as.numeric(substr(land$SYEAR,6,9))
	land =land[order(land$YR),]


		d1 = data.frame(YEAR = land$YR, LANDINGS = land[,paste0("LFA",p$lfas[i])])
		d2 = subset(CPUEindex,LFA==p$lfas,c("LFA","YEAR","mu"))
		names(d2)[3]="CPUE"

		d2  = merge(data.frame(LFA=d2$LFA[1],YEAR=min(d2$YEAR):max(d2$YEAR)),d2,all.x=T)

		fishData = merge(d2,d1) 
		fishData$EFFORT2 = fishData$LANDINGS * 1000 / fishData$CPUE

	
		tiff("C:/Users/Howsevj/Documents/LFA35_Update/35landings.tiff", width=8,height=6, units="in",res=300)

	par(mar=c(5.1, 4.1, 4.1, 5.1),las=1)
			
  plot(fishData$YEAR,fishData$LANDINGS,xlab='Year',ylab='Landings(t)',type='h',ylim=c(0,max(fishData$LANDINGS)*1.2),pch=15,col='grey',lwd=10,lend=3)
		lines(fishData$YEAR[nrow(fishData)],fishData$LANDINGS[nrow(fishData)],type='h',pch=21,col='steelblue4',lwd=10,lend=3)
		par(new=T)
		plot(fishData$YEAR,fishData$EFFORT2/1000,ylab='',xlab='', type='b', pch=16, axes=F,ylim=c(0,max(fishData$EFFORT2/1000,na.rm=T)))
	points(fishData$YEAR[nrow(fishData)],fishData$EFFORT2[nrow(fishData)]/1000, type='b', pch=24,bg='black')
		axis(4)
		mtext("Effort ('000s Trap Hauls)", 4, 3.5, outer = F,las=0)	
dev.off()		
```




```{r Stratified Analysis Commercial,include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE }
require(bio.survey)
require(bio.lobster)

p = bio.lobster::load.environment()
p$libs = NULL
ff = "LFA35-38Assessment"
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
dir.create(fpf1,showWarnings=F)
dir.create(fp1,showWarnings=F)

p$yrs = 1970:2019
p1 = p


stratifiedAnalysesCommercial = function( p=p1, survey,lfa, fpf = fpf1, fp = fp1,f=ff,wd=10,ht=8){
    p$series =c('summer')# p$series =c('georges');p$series =c('fall')
    p$years.to.estimate = p$yrs
    p$length.based = T
    p$by.sex = T
    p$size.class = c(83,300)
    p$sex = c(1,2)
    p$bootstrapped.ci=T
    p$strata.files.return=F
    p$vessel.correction.fixed=1.2
    p$strat = NULL
    p$clusters = c( rep( "localhost", 7) )
    p$strata.efficiencies = F
    p = make.list(list(yrs=p$years.to.estimate),Y=p)
    p$define.by.polygons = T
    p$lobster.subunits=F
    p$area = lfa
    p$reweight.strata = T #this subsets 

    aout= dfo.rv.analysis(DS='stratified.estimates.redo',p=p)
    write.csv(aout,file=file.path(fpf, paste(lfa,'DFOCommB.csv')))

      p$add.reference.lines = F
    p$time.series.start.year = p$years.to.estimate[1]
    p$time.series.end.year = p$years.to.estimate[length(p$years.to.estimate)]
    p$metric = 'weights' #weights
    p$measure = 'stratified.total' #'stratified.total'
    p$figure.title = ""
    p$reference.measure = 'median' # mean, geomean
    p$file.name =  file.path(f,paste(lfa,'DFOrestratifiedtotalweightscommercial.png',sep=""))

    p$y.maximum = NULL # NULL # if ymax is too high for one year
    p$show.truncated.weights = F #if using ymax and want to show the weights that are cut off as values on figure

    p$legend = FALSE
    p$running.median = T
    p$running.length = 3
    p$running.mean = F #can only have rmedian or rmean
     p$error.polygon=F
    p$error.bars=T
    require(bio.lobster)
     p$ylim2 = NULL
    xx = aggregate(ObsLobs~yr,data=aout,FUN=sum)
    names(xx) =c('x','y')
     p$ylim=NULL
     ref.out= figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
     return(aout)
  }



aout = stratifiedAnalysesCommercial(survey='DFO',lfa='LFA35-38')
write.csv(aout,file.path(fp1,'LFA3538CommercialB.csv'))
```


```{r Stratified Analysis,include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE}
require(bio.survey)
require(bio.lobster)
require(lubridate)
require(devtools)
require(bio.utilities)
require(PBSmapping)
la()
p = bio.lobster::load.environment()
p$libs = NULL
ff = "LFA35-38Assessment"
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
dir.create(fpf1,showWarnings=F)
dir.create(fp1,showWarnings=F)
p1 = p
p1$yrs = 1970:2019

stratifiedAnalyses = function(p=p1, survey,lfa, fpf = fpf1, fp = fp1,f=ff,wd=10,ht=8){
                p$series =c('summer')
                p$define.by.polygons = T
                p$lobster.subunits=F
                p$area = lfa
                p$years.to.estimate = p$yrs[-1]
                p$length.based = F
                p$by.sex = F
                p$bootstrapped.ci=T
                p$strata.files.return=F
                p$vessel.correction.fixed=1.2
                p$strat = NULL
                p$clusters = c( rep( "localhost", 7) )
                p$strata.efficiencies = F
                p = make.list(list(yrs=p$years.to.estimate),Y=p)
                p$reweight.strata = T #this subsets 
            

      aout= dfo.rv.analysis(DS='stratified.estimates.redo',p=p)
      write.csv(aout,file=file.path(fpf, paste(lfa,'DFOtotalabund.csv')))


                              p$add.reference.lines = F
                              p$time.series.start.year = p$years.to.estimate[1]
                              p$time.series.end.year = p$years.to.estimate[length(p$years.to.estimate)]
                              p$metric = 'numbers' #weights
                              p$measure = 'stratified.mean' #'stratified.total'
                              p$figure.title = ""
                              p$reference.measure = 'median' # mean, geomean
                              p$file.name = file.path(f,paste(lfa,'DFOrestratifiednumbers.png',sep=""))

                          p$y.maximum = NULL # NULL # if ymax is too high for one year
                        p$show.truncated.numbers = F #if using ymax and want to show the numbers that are cut off as values on figure

                                p$legend = FALSE
                                p$running.median = T
                                p$running.length = 3
                                p$running.mean = F #can only have rmedian or rmean
                               p$error.polygon=F
                              p$error.bars=T

                               p$ylim=NULL
                      if(lfa == 'LFA35-38') p$ylim=c(0,150)
                       p$box=T
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiednumbersNOY.png', sep=""))
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)

                       p$box=T
                        p$ylim=NULL
                       p$metric = 'weights'
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiedweightsNOY.png',sep=""))
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)

                       p$box=NULL
                       p$ylim=NULL
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiedDWAO.png',sep=""))
                       p$metric = 'dwao'
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
                       
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiedgini.png',sep=""))
                       p$metric = 'gini'
                       p$ylim =c(0,1)
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
                       p$ylim = NULL
                       return(aout)
        }

    
a = stratifiedAnalyses(survey='DFO',lfa='LFA35-38')

```


```{r Stratified Analysis Recruits,include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE}
require(bio.survey)
require(bio.lobster)
require(lubridate)
require(devtools)
require(bio.utilities)
require(PBSmapping)
la()
p = bio.lobster::load.environment()
p$libs = NULL
ff = "LFA35-38Assessment"
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
dir.create(fpf1,showWarnings=F)
dir.create(fp1,showWarnings=F)
p1 = p
p1$yrs = 1970:2019

stratifiedAnalysesRecruits = function(p=p1, survey,lfa, fpf = fpf1, fp = fp1,f=ff,wd=10,ht=8){
      p$series =c('summer')# p$series =c('georges');p$series =c('fall')
      p$area = lfa
      p$years.to.estimate = p$yrs
      p$length.based = T
      p$by.sex = T
      p$size.class = c(70,82)
      p$sex = c(1,2)
      p$bootstrapped.ci=T
      p$strata.files.return=F
      p$vessel.correction.fixed=1.2
      p$strat = NULL
      p$clusters = c( rep( "localhost", 7) )
      p$strata.efficiencies = F
      p = make.list(list(yrs=p$years.to.estimate),Y=p)
      p$define.by.polygons = T
      p$lobster.subunits=F
      p$reweight.strata = T #this subsets 
  
    aout= dfo.rv.analysis(DS='stratified.estimates.redo',p=p)
     write.csv(aout,file=file.path(fpf, paste(lfa,'DFOrecruits.csv')))

                        p$add.reference.lines = F
                              p$time.series.start.year = p$years.to.estimate[1]
                              p$time.series.end.year = p$years.to.estimate[length(p$years.to.estimate)]
                              p$metric = 'numbers' #numbers
                              p$measure = 'stratified.mean' #'stratified.total'
                              p$figure.title = ""
                              p$reference.measure = 'median' # mean, geomean
                              p$file.name =  file.path(f,paste(lfa,'DFOrestratifiednumbersrecruits.png',sep=""))
                          p$y.maximum = NULL # NULL # if ymax is too high for one year
                        p$show.truncated.numbers = F #if using ymax and want to show the numbers that are cut off as values on figure
                                p$legend = FALSE
                                p$running.median = T
                                p$running.length = 3
                                p$running.mean = F #can only have rmedian or rmean
                               p$error.polygon=F
                              p$error.bars=T
                      if(lfa == 'LFA35-38') p$ylim = c(0,80)
                      p$file.name =  file.path(f,paste(lfa,'NOYDFOrestratifiednumbersrecruits.png',sep=""))
                      ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
            }

stratifiedAnalysesRecruits(survey='DFO',lfa='LFA35-38')



```



```{r Landings Survey Relative F, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE}

require(bio.lobster)
require(PBSmapping)

a = lobster.db('annual.landings')
b = lobster.db('seasonal.landings')
ff = "LFA35-38Assessment"

b$YR = substr(b$SYEAR,6,9)
a = subset(a,YR<1976)
b = subset(b,YR>1975 & YR<=2019)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)

LFA = c('LFA35','LFA36','LFA38')
for(i in LFA){
		aa = a[,c('YR',i)]
		bb = b[,c('YR',i)]
		aa = (rbind(aa,bb))
		aa = aa[order(aa$YR),]
		file.name = paste('Landings',i,'.png',sep="")
		png(file=file.path(fpf1,'LandingsL3538.png'),units='in',width=15,height=12,pointsize=18, res=300,type='cairo')
		plot(aa$YR,aa[,i],type='h',lwd=4,col='black',xlab='Year',ylab='Landings (t)')
		lines(aa$YR,runmed(aa[,i],3),col='salmon',lwd=3)
		dev.off()
		}


###lfa 35-38

df2 = read.csv(file.path(fpf1,'LFA35-38 DFOCommB.csv'))
df =  read.csv(file.path(fpf1,'LFA35-38 DFOtotalabund.csv'))
df = subset(df,yr<1999)
df2 = subset(df2,yr>1998)
df = as.data.frame(rbind(df,df2))
df = df[,c('yr','w.Yst','w.ci.Yst.l','w.ci.Yst.u')] #proportion of total weight that is commercial
df$w.Yst[which(df$yr<1999)] <- df$w.Yst[which(df$yr<1999)]*0.746
df$w.ci.Yst.l[which(df$yr<1999)] <- df$w.ci.Yst.l[which(df$yr<1999)]*0.746
df$w.ci.Yst.u[which(df$yr<1999)] <- df$w.ci.Yst.u[which(df$yr<1999)]*0.746



 	 #png(file=file.path(fpf1,'LFA35-38CommBDFOextended.png'),units='in',width=10,height=8,pointsize=18, res=300,type='cairo')
 	 with(df,plot(yr,w.Yst,pch=1,xlab='Year',ylab='Commerical Biomass (t)',ylim=c(0,9500)))
	 with(df,arrows(yr,y0=w.ci.Yst.u,y1=w.ci.Yst.l, length=0))
	 with(subset(df,yr>1998),points(yr,w.Yst,pch=16))
	 xx = rmed(df$yr,df$w.Yst)
	 xx = as.data.frame(do.call(cbind,xx))
	 with(subset(xx,yr<1999),lines(yr,x,col='salmon',lwd=1))
	 with(subset(xx,yr>1998),lines(yr,x,col='salmon',lwd=3))
	 dev.off()
	 
a$L3538 = rowSums(a[,12:14])
b$L3538 = rowSums(b[,4:6])
c358 = as.data.frame(rbind(a[,c('YR','L3538')],b[,c('YR','L3538')]))
names(c358)[1] = 'yr'
df  =merge(df,c358)
df$rL = df$L3538/(df$w.ci.Yst.l+df$L3538)
df$rU =df$L3538/ (df$w.ci.Yst.u+df$L3538)
df$rM = df$L3538/(df$w.Yst+df$L3538)


 	 png(file=file.path(fpf1,'LFA3538RelFDFO.png'),units='in',width=10,height=8,pointsize=18, res=300,type='cairo')
		plot(df$yr,df$rM,type='p',pch=16,col='black',xlab='Year',ylab='Relative F')
		arrows(df$yr,y0 = df$rL,y1 = df$rU,length=0)
		with(rmed(df$yr,df$rM),lines(yr,x,col='salmon',lwd=3))
dev.off()

```
#DFO RV Survey Commercial Biomass and Recruit Abundance

```{r Commercial Biomass,include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE, fig.cap="Time series of DFO RV survey trends for LFA 35-38 commercial biomass. Values prior to 1999 were derived using the mean proportion of commercial to total biomass between 1999 and 2018 (0.746)."}


```

```{r Recruit Abundance,include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE, fig.cap="Time series of DFO RV survey trends for LFA 35-38 recruit abundance. Red line represents the 3 year running median."}

```

```{r Relative F,include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE, fig.cap="Relative fishing mortality from the DFO RV survey commercial biomass estimates  and the landings in LFA 35-38."}

```

