  
---
title: "Bycatch Data Submission"
author: "Lobster Group, Science Branch "
output: tint::tintPdf
---


**1. Data was collected**
```{r echo =FALSE, message =FALSE, error = FALSE, warning = FALSE, results='asis' , fig-main, fig.width=10, fig.height=12,  cache=TRUE,}

    dat$Y<- convert.dd.dddd(dat$LATDDMM, format="dec.deg")
    dat$X<- convert.dd.dddd(dat$LONGDDMM, format="dec.deg")*-1
    dat$EID<- 1:nrow(dat)

    xlim = c(min(dat$X)-0.5, max(dat$X)+0.5)
    ylim = c(min(dat$Y)-0.5, max(dat$Y)+0.5)

    x11(10,10)
    LobsterMap(ylim=ylim ,xlim=xlim, labels = 'lfa')
    attr(dat,"projection") <- "LL"
    addPoints(dat, pch = 21, cex=1.5, bg='red')

    ll<-unique(dat$COMAREA_ID)
    lll<-ifelse(nchar(ll)==3,3,4)
    ll<-substr(ll,2,lll)
    title(paste0("LFA", ll))
```
  
**2. What group entered the data and the name of the data file that was imported**
```{r echo=FALSE, results='asis'}
    cat(paste("<br> <ul>", unique(dat$OWNER_GROUP),unique(dat$TRIP),"</br> </ul>"))  
    if(!grepl(unique(dat$TRIP),nm)){
          print(paste("<br><ul> \textcolor{red}{Trip File Name does not match TRIP column in Data Sheet. Check ",Fi[i],"} </br></ul>",sep=" "))
      }
      
``` 


 **3. Do the coordinates match the Grids and LFA?**
```{r  echo=FALSE, results = 'asis'}
    dat$COMAREA_ID[which(dat$COMAREA_ID =="L31A")]<-311
    dat$COMAREA_ID[which(dat$COMAREA_ID =="L31B")]<-312

    ll=unique(dat$COMAREA_ID)

    if(!unique(dat$COMAREA_ID) %in% c(311,312)){
        ll<-unique(dat$COMAREA_ID)
        lll<-ifelse(nchar(ll)==3,3,4)
        ll<-substr(ll,2,lll)
    }

    LFAs<-subset(LFAS, PID ==ll )
   LFAgrid<-subset(LFAGrid, PID == ll)


    #Convert to Decimal Degrees
    dat$Y<-convert.dd.dddd(dat$LATDDMM, format="dec.deg")
    dat$X<-(convert.dd.dddd(dat$LONGDDMM, format="dec.deg")*-1)
     
    dat$EID = 1:nrow(dat)

    #Check LFA Locations
    locCheck<-findPolys(dat, LFAs)
    if(is.null(locCheck)){
        cat(paste0("<br> <ul> \textcolor{red}{ERROR:Wrong LFA identified} </br> </ul>"))
          }else{      
        missingLoc<-merge(dat, locCheck, all.x = T, by = "EID")
        missedLoc<-subset(missingLoc, is.na(PID))

        if(nrow(missedLoc)>0){ 
            print(paste0("<br> <ul> \textcolor{red}{ERROR:LFA locations missing or incorrect. Check line(s)",paste(as.numeric(row.names(missedLoc))+1, collapse=" , "),"}", sep=" ", "</br> </ul>"))
          } else {print("<br> <ul> <center> All locations match LFA </br> </ul> </center>")
        }
    #Check Grid Locations
    glocCheck<-findPolys(dat, LFAgrid)

    #Find missing grid locations - or locations that don't pair with corresponding grids
        missingGLoc<-merge(dat, glocCheck, all.x = T, by = "EID")
        missedGLoc<-subset(missingGLoc, is.na(PID))

    if(nrow(missedGLoc)>0){ 
          print(paste0("<br> <ul> \textcolor{red}{ERROR: Grid locations missing or incorrect. Check line(s) ", paste(as.numeric(row.names(missedGLoc))+1,collapse=","), "}", sep=" ", "</br> </ul>"))
        } else {cat("<br> <ul> All locations match grids </br> </ul>")
      }
    }
      
```
 
**4. Are males berried or notched?**
```{r echo=FALSE, results = 'asis'}
    dat$SEXCHECK<-NA
    dat$SEXCHECK<- ifelse(dat$SPECCD_ID ==2550 & dat$SEXCD_ID == 1 & !is.na(dat$VNOTCH) , 1, dat$SEXCHECK)
      
    notchcheck<-any(dat$SEXCHECK == 1, na.rm = T)
    notchmales<-which(dat$SEXCHECK == 1)


    if(notchcheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR:Check SEX and NOTCH. Errors on line(s) ",paste(notchmales+1, collapse=" , "), "}", sep=" ", "</br> </ul>"))
      }else {cat("<br> <ul> No Males are V-notched </br> </ul>")
      }
    dat$SEXCHECK <- NULL
```

**5. Is the length between a reasonable size 30-250mm?**
```{r echo=FALSE, results = 'asis'}

    dat$CLCHECK<-NA
    dat$CLCHECK<- ifelse(dat$SPECCD_ID == 2550 & (dat$FISH_LENGTH <=30 | dat$FISH_LENGTH >=250), 1, dat$CLCHECK)

    sizecheck<-any(dat$CLCHECK ==1, na.rm=T)
    wrongsize<-which(dat$CLCHECK==1)

    if(sizecheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR: Sizes are not reasonable. Check for errors on line(s) ", paste(wrongsize+1, collapse=" , "), "}", sep= " ", "</br> </ul>"))
    }else{cat("<br> <ul> Carapace sizes reasonable </br> </ul>")}

    dat$CLCHECK <- NULL 
      
```


**6. Is Length, Shell, Disease, Sex and Condition recorded for all Lobster?** 
```{r echo=FALSE, results = 'asis'}

    dat$LENGTHDAT<-NA
    dat$LENGTHDAT<-ifelse(dat$SPECCD_ID == 2550 & is.na(dat$FISH_LENGTH), 1, dat$LENGTHDAT)
    lengthcheck<-any(dat$LENGTHDAT == 1, na.rm=T)
    nolength<-which(dat$LENGTHDAT==1)
    if(lengthcheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR: Missing information for carapace length. Check line(s) ", paste(nolength+1, collapse=" , "), "}", sep=" ",  "</br> </ul>"))
    }else{ cat("<br> <ul> Carapace lengths recorded </br> </ul> ")
    }

    dat$SHELLDAT<-NA
    dat$SHELLDAT<-ifelse(dat$SPECCD_ID == 2550 & is.na(dat$SHELL), 1, dat$SHELLDAT)
    shellcheck<-any(dat$SHELLDAT == 1, na.rm=T)
    noshell<-which(dat$SHELLDAT==1)
    if(shellcheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR: Missing information for shell hardness. Check line(s) ", paste(noshell+1, collapse=" , "),"}", sep = " ", "</br> </ul>"))
    }else{ cat("<br> <ul> Shell hardness recorded </br> </ul>")
    }

    dat$SICDAT<-NA
    dat$SICDAT<-ifelse(dat$SPECCD_ID == 2550 & is.na(dat$DISEASE), 1, dat$SICDAT)
    siccheck<-any(dat$SICDAT == 1, na.rm=T)
    nosic<-which(dat$SICDAT==1)
    if(siccheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR: Missing information for disease. Check line(s) ", paste(nosic+1, collapse=" , "), "}", sep = " ", "</br> </ul>"))
    }else{ cat("<br> <ul> Disease status recorded </br> </ul>")
    }


    dat$SEXDAT<-NA
    dat$SEXDAT<-ifelse(dat$SPECCD_ID == 2550 & is.na(dat$SEXCD_ID), 1, dat$SEXDAT)
    sexcheck<-any(dat$SEXDAT == 1, na.rm=T)
    nosex<-which(dat$SEXDAT==1)
    if(sexcheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR: Missing information for sex. Check line(s) ", paste(nosex+1, collapse=" , "), "}",sep=" ", "</br> </ul>"))
    }else{ cat("<br> <ul> Sex recorded </br> </ul>")
    }

    dat$CONDAT<-NA
    dat$CONDAT<-ifelse(dat$SPECCD_ID == 2550 & is.na(dat$CONDITION_CD), 1, dat$CONDAT)
    concheck<-any(dat$CONDAT == 1, na.rm=T)
    nocon<-which(dat$CONDAT==1)
    if(concheck == T){
      print(paste0("<br> <ul> \textcolor{red}{ERROR: Missing information for condition. Check line(s) ", paste(nocon+1, collapse=" , "), "}", sep=" ", "</br> </ul>"))
    }else{ cat("<br> <ul> Condition recorded </br> </ul>")
    }
      
```


**7. Is the depth less than a maximum fishing depth ?**
```{r echo=FALSE, results = 'asis'}
      dat$DEPTHCHECK<-NA
      dat$DEPTHCHECK<-ifelse(dat$DEPTH <= 300,0,1)

      deepcheck<-any(dat$DEPTHCHECK ==1,na.rm=T)
      baddepth<-which(dat$DEPTHCHECK==1)

      if(deepcheck==T){
        print(paste0("<br> <ul> \textcolor{red}{ERROR: Depths exceed a maximum fishing depth. Check line(s) ", paste(baddepth+1, collapse=" , "), "}", sep = " ", "</br> </ul>"))
      }else{cat("<br> <ul> Depths within reasonable range </br> </ul>")}

      dat$DEPTHCHECK <- NULL
        
```


**8. Were any lobster kept below the MLS for the given LFA?**
      ```{r echo= FALSE, results = 'asis'}
      MLS=82.5
      dat$smallretained <-0

      if(unique(dat$COMAREA_ID) %in% c("L28","L29")){
        MLS=84 }

      dat$smallretained[which(dat$SPECCD_ID ==2550 & dat$FISH_LENGTH < MLS & dat$KEPT == 1)]<-1

      if(any(dat$smallretained ==1) ){
        ij<-which(dat$smallretained==1)
        print(paste0("<br> <ul> \textcolor{red}{ERROR:Lobsters kept below MLS. Check line(s) ", paste(ij+1, collapse=" , "), "}",sep = " ", "</br> </ul>"))
      }else{cat("<br> <ul> No small lobsters retained </br> </ul>")}

      dat$smallretained = NULL
  
```


**9. Were female between 114-124mm kept in LFA 31a?**
```{r echo = FALSE, results = 'asis'}
        dat$window<- 0
        dat$window<- ifelse(dat$COMAREA_ID %in% c(311) & dat$SEXCD_ID %in% c(2,3) & dat$SPECCD_ID == 2550  & dat$FISH_LENGTH %in% c(114:124) & dat$KEPT %in% c(1), 1, 0) 
        if(any(dat$window == 1)){
              ko<-which(dat$window ==1)
              print(paste0("<br> <ul> \textcolor{red}{ERROR:Lobsters kept between 114-124mm in 31a. Check line(s)", paste(ko+1, collapse=" , "), "}", sep = " ", "</br> </ul>"))
            } else {
              cat("<br> <ul> No window restricted lobsters were kept </br> </ul>")
      }
       dat$window = NULL
```


**10. Were v-notched lobster retained in prohibited LFAs?**
```{r echo = FALSE, results = 'asis'}
            dat$keepnotch<-0
       if(any(is.na(dat$KEPT))) dat$KEPT[which(is.na(dat$KEPT))] <- 0
              dat$keepnotch<- ifelse(dat$COMAREA_ID %ni% c("L27","L31A") & !is.na(dat$VNOTCH) & dat$KEPT == 1,1,0)
       if(any(dat$keepnotch ==1)){
              pl<- which(dat$keepnotch==1)
              print(paste0("<br> <ul> \textcolor{red}{ERROR:V-notched lobsters retained in prohibited LFA. Check line(s) ", paste(pl+1, collapse=" , "), "}", sep = " ", "</br> </ul>"))
          }else{
              cat("<br> <ul> No vnotched lobster retained in prohibited area </br> </ul>")
          }
        dat$keepnotch =NULL
```


**11. Is abundance greater than 1 for anything other than whelks, starfish, urchins?**
```{r echo = FALSE, results = 'asis'}
          dat$AMOUNT<-NA
          dat$AMOUNT<- ifelse((dat$SPECSCD_ID <= 1000 | dat$SPECCD_ID %in% c(2550,2513,2511,2531,2520,2526)) & (dat$ABUNDANCE > 1), 1, 0)

          amountcheck<- any(dat$AMOUNT ==1, na.rm=T)
          wrongamount<-which(dat$AMOUNT==1)
        if(amountcheck ==T){ 
              print(paste0("<br> <ul> \textcolor{red}{ERROR:Check Abundance. Error on line(s)", paste(wrongamount+1, collapse=" , "), "}", sep = " ", "</br> </ul>"))
            } else {
              cat("<br> <ul> Abundance is correct </br> </ul>") 
            }
        dat$AMOUNT<-NULL
```
 
 
**12. If species is not a berried lobster does it have clutch and egg categories filled out?**
```{r echo = FALSE, results = 'asis'}
          dat$BERRIED<-NA
          dat$BERRIED<- ifelse(dat$SPECCD_ID ==2550 & dat$SEXCD_ID %in% c(1,2) & c(!is.na(dat$CLUTCH) | !is.na(dat$EGG_STAGE)),1,0)
          notberried<- any(dat$BERRIED == 1, na.rm=T)
          noberries<-which(dat$BERRIED==1)
  if(notberried ==T){
          print(paste0("<br> <ul> \textcolor{red}{ERROR:Check SEX, CLUTCH, an/or EGG STAGE. Error on line(s) ", paste(noberries+1, collapse=" , "), "}", sep = " ", "</br> </ul>"))
          } else {
            cat("<br> <ul> Only berried females have egg and clutch information  </br> </ul>")
          }
    dat$BERRIED<-NULL
```


**13. If a lobster is berried is egg stage and clutch size recorded?**
```{r echo = FALSE, results = 'asis'}
          dat$BERRIES<-NA
          dat$BERRIES<- ifelse(dat$SPECCD_ID ==2550 & dat$SEXCD_ID== 3 & c(is.na(dat$CLUTCH) | is.na(dat$EGG_STAGE)),1,0)
          berried<- any(dat$BERRIES == 1, na.rm=T)
          noberried<-which(dat$BERRIES==1)
    if(berried ==T){
            print(paste0("<br> <ul> \textcolor{red}{ERROR:Check CLUTCH and/or EGG STAGE. Error on line(s) ", paste(noberried+1, collapse=" , "), "}", sep= " ", "</br> </ul>"))
          } else {
          cat("<br> <ul> Berried females have egg and clutch information </br> </ul>")
        }
dat$BERRIES<-NULL
```


**14. Do all crab species entries have length, sex and condition recorded?**
```{r echo = FALSE, results = 'asis'}
          dat$CRABBIE<-NA
          dat$CRABBIE<- ifelse(dat$SPECCD_ID %in% c(2513,2511,2531,2520,2526) & (dat$FISH_LENGTH < 1 | is.na(dat$FISH_LENGTH)) & (!dat$SEXCD_ID %in% c(1,2,3) | is.na(dat$SEXCD_ID)), 1, 0)
          crabinfo<- any(dat$CRABBIE ==1, na.rm=T)
          nocrab<- which(dat$CRABBIE ==1)
    if(crabinfo ==T){
                print(paste0("<br> <ul> <center> \textcolor{red}{ERROR:Check SEX and/or FISH LENGTH. Error on line(s) ", paste(nocrab+1, collapse=" , "), "}", sep= " ", "</br> </ul> </center>"))
          } else {
            cat("<br> <ul> All crabs have length and sex recorded </br> </ul>")
          }
      dat$CRABBIE<- NULL 

      dat$CCONDAT<-NA
      dat$CCONDAT<-ifelse(dat$SPECCD_ID %in% c(2513,2511,2531,2520,2526)  & is.na(dat$CONDITION_CD), 1, dat$CCONDAT)
      cconcheck<-any(dat$CCONDAT == 1, na.rm=T)
      noccon<-which(dat$CCONDAT==1)
    if(cconcheck == T){
              print(paste0("<br> <ul> \textcolor{red}{ERROR: Missing information for condition. Check line(s) ", paste(nocon+1, collapse=" , "), "}", sep= " ","</br> </ul>"))
      } else { 
        cat("<br> <ul> Condition recorded </br> </ul>")
    }
    dat$CCONDAT <- NULL
```


**15. Does anything other than a lobster have the notch category filled out?**
```{r echo = FALSE, results = 'asis'}
        dat$NOTCHED<- NA
        dat$NOTCHED<- ifelse(!dat$SPECCD_ID==2550 & !is.na(dat$VNOTCH),1,0)
        nonotch<- any(dat$NOTCHED==1, na.rm=T)
        notchless<- which(dat$NOTCHED ==1)
    if(nonotch ==T){
              print(paste0("<br> <ul> \textcolor{red}{ERROR:Check NOTCH on line(s) ", paste(notchless+1, collapse=" , "),"}", sep= " ", "</br> </ul>"))
        } else {
          cat("<br> <ul> Only lobster have notches </br> </ul>")
        }
    dat$NOTCHED<-NULL
```


**16. Do all Fish species entries have length, condition recorded?**
```{r echo = FALSE, results = 'asis'}
      dat$FISHY<-NA
      dat$FISHY<- ifelse(dat$SPECCD_ID <1000 & c(dat$FISH_LENGTH < 1 | is.na(dat$FISH_LENGTH)), 1, 0)
      fishinfo<- any(dat$FISHY ==1, na.rm=T)
        nofish<- which(dat$FISHY ==1)
  if(fishinfo ==T){
        print(paste0("<br> <ul> \textcolor{red}{ERROR: Check FISH LENGTH on line(s) ", paste(nofish+1, collapse=" , "),"}", "</br> </ul>",sep=" "))
      } else {
      cat("<br> <ul> All fish have length recorded </br> </ul>")
    }
    dat$FISHY<- NULL 

      dat$CCONDAT<-NA
      dat$CCONDAT<-ifelse(dat$SPECCD_ID %in% c(1:1000,2513,2511,2531,2520,2526)  & is.na(dat$CONDITION_CD), 1, dat$CCONDAT)
      cconcheck<-any(dat$CCONDAT == 1, na.rm=T)
      noccon<-which(dat$CCONDAT==1)
    if(cconcheck == T){
        cat(paste0(" \textcolor{red}{ERROR: Missing information for CONDITION. Check line(s) ", paste(noccon+1, collapse=" , "),"}", sep=" "))
        } else {
         cat("<br> <ul> All fish have condition recorded </br> </ul>")
      }
      dat$CCONDAT <- NULL
  
```

**17. What is the trap sampling frequency per string?**
```{r echo = FALSE, results = 'asis'}
trapnum<-as.data.frame(unique(cbind(dat$SET_NO,dat$TRAP_NO)))
trapnum<-trapnum[order(trapnum$V1,trapnum$V2),]
trapfreq<-aggregate(V2~V1, data=trapnum,FUN = function(x) mean(diff(x)))
trapfreq$test<-round(trapfreq$V2,0)
names(trapfreq)[1:2]<-c("String", "Frequency")
irrfreq<-which(trapfreq$Frequency != trapfreq$test)
trapfreq$test <-NULL
if(length(irrfreq)> 0){
  print(paste0("<br> <ul> \textcolor{red}{ERROR:Irregular trap sampling frequencies in string(s) ", trapfreq[irrfreq,"String"], "}", sep=" ", "</br> </ul>" ))
}
if(length(irrfreq)<nrow(trapfreq)){
  cat("<br> <ul> Trap sampling frequencies by string </br> </ul>")
  if(length(irrfreq)>=0)cat(paste0("<ul> One in ",  unique(trapfreq[-irrfreq,'Frequency']), " traps sampled per string </ul>"))
  }
  
```


**18. What is the total number of traps sampled and total traps hauled**
```{r echo= FALSE, results = 'asis'}

sampled<-length(unique(paste(dat$TRAP_NO,dat$SET_NO,sep="-")))

if(length(sampled)>0){
  cat(paste0("<br> <ul> Total number of traps sampled: ", sampled, "</br> </ul>"))
}

totaltrap<-as.data.frame(unique(cbind(dat$SET_NO, dat$NUM_HOOK_HAUL)))

if(sum(totaltrap$V2, na.rm = T)< sampled){
  print("<br> <ul> \textcolor{red}{ERROR: Check SET NO and NUM HOOK HAUL numbers do not add up to total traps sampled.}, </br> </ul>")
} else {cat(paste0("<br> <ul> Total number of traps hauled: ", sum(totaltrap$V2),"</br> </ul>"))
}
  
```


**19. What was the total number of lobsters and other species sampled?**
```{r echo=FALSE, results = 'asis'}
numlob<-length(which(dat$SPECCD_ID==2550))

if(length(numlob>0)){
  cat(paste0("<br> <ul> Total number of lobsters sampled: ", numlob, "</br> </ul>"))
      
}

numother<-length(which(dat$SPECCD_ID != 2550))

if(length(numother>0)){
  cat(paste0("<br> <ul> Total number of non lobsters sampled: ", numother, "</br> </ul>"))
}


numother<-length(unique(dat$SPECCD_ID))

if(length(numother>0)){
  cat(paste0("<br> <ul> Total number of species: ", numother, "</br> </ul>"))
  }
``` 