---
title: "Lobster Fishery Viability Analysis"
author: "DFO Precautionary Approach Framework"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cosmo
---

# Overview

This document demonstrates a DFO-style precautionary approach assessment using:

- Seasonal temperatures
- Population abundance estimates
- Fishing effort (days fished)
- Effort-dependent depletion
- Biomass reference points
- Removal reference (FMSY/2)
- Stochastic projections

# Required Packages

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(minpack.lm)
library(knitr)
library(scales)

theme_set(theme_bw())
```

# Generate Example Dataset

```{r generate-data}

set.seed(123)

years <- 2000:2025
n <- length(years)

SpringTemp <- round(runif(n, 2, 11), 2)
SummerTemp <- round(runif(n, 7, 14), 2)
FallTemp   <- round(runif(n, 10, 16), 2)

DaysFished <- round(runif(n, 45, 55))

ER <- pmin(
  pmax(
    rnorm(
      n,
      mean = 0.25,
      sd = 0.05
    ),
    0.10
  ),
  0.50
)

Population <- numeric(n)

Population[1] <- 12000

for(i in 1:(n-1)) {

  r_true <- 0.8
  K_true <- 30000

  growth <-
    r_true *
    Population[i] *
    (1 - Population[i] / K_true)

  catch_now <-
    Population[i] * ER[i]

  Population[i + 1] <-
    Population[i] +
    growth -
    catch_now +
    rnorm(1, 0, 500)

  Population[i + 1] <-
    min(
      max(Population[i + 1], 10000),
      30000
    )
}

peak_effect <-
  8000 *
  exp(
    -((years - 2015)^2) /
      (2 * 4^2)
  )

Population <-
  Population + peak_effect

Population <-
  pmin(Population, 30000)

Population <- round(Population)

Catch <- round(
  Population * ER
)

lobster_data <- data.frame(
  Year = years,
  SpringTemp,
  SummerTemp,
  FallTemp,
  Population,
  DaysFished,
  Catch
)

write.csv(
  lobster_data,
  "lobster_data.csv",
  row.names = FALSE
)

head(lobster_data)

```

# Load Data

```{r}
dat <- read.csv("lobster_data.csv")

head(dat)

summary(dat)
```

# Create Thermal Index

The thermal index summarizes seasonal environmental conditions.

```{r}

dat <- dat %>%
  mutate(
    TempIndex =
      0.4 * SpringTemp +
      0.4 * SummerTemp +
      0.2 * FallTemp
  )

summary(dat$TempIndex)

```

# CPUE

```{r}

dat <- dat %>%
  mutate(
    CPUE = Catch / DaysFished
  )

ggplot(dat, aes(Year, CPUE)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  labs(
    y = "CPUE",
    title = "Catch Per Unit Effort"
  )

```

# Effort Depletion Model

Assume effective catchability declines as fishing effort increases.

\[
q_{eff} = q e^{-dE}
\]

where:

- \(q\) = baseline catchability
- \(d\) = depletion parameter
- \(E\) = fishing effort in days

```{r}

depletion_model <- nlsLM(

  CPUE ~
    q *
    Population *
    exp(-d * DaysFished),

  data = dat,

  start = list(
    q = mean(dat$CPUE) /
      mean(dat$Population),
    d = 0.01
  )

)

summary(depletion_model)

q_hat <- coef(depletion_model)["q"]
d_hat <- coef(depletion_model)["d"]

```

# Estimate Fishing Mortality

Assume:

\[
F_t = q_t E_t
\]

```{r}

dat <- dat %>%
  mutate(
    F_est =
      q_hat *
      exp(-d_hat * DaysFished) *
      DaysFished
  )

summary(dat$F_est)

```

# Population Dynamics Model

Population dynamics:

\[
B_{t+1}
=
B_t
+
r_t B_t
\left(1-\frac{B_t}{K}\right)
-
F_t B_t
\]

Temperature-dependent productivity:

\[
r_t = r e^{\beta T_t}
\]

```{r}

negLL <- function(par, dat){

  r <- exp(par[1])
  K <- exp(par[2])
  beta <- par[3]

  n <- nrow(dat)

  B <- numeric(n)
  B[1] <- dat$Population[1]

  nll <- 0

  for(i in 1:(n - 1)) {

    Fnow <-
      q_hat *
      exp(-d_hat * dat$DaysFished[i]) *
      dat$DaysFished[i]

    r_eff <-
      r *
      exp(beta * dat$TempIndex[i])

    pred <-
      B[i] +
      r_eff *
      B[i] *
      (1 - B[i] / K) -
      Fnow *
      B[i]

    pred <- max(pred, 1)

    B[i + 1] <- pred

    nll <- nll -
      dnorm(
        dat$Population[i + 1],
        mean = pred,
        sd = max(0.10 * pred, 1),
        log = TRUE
      )
  }

  nll
}

```

# Estimate Parameters

```{r}

fit <- optim(

  par = c(
    log(0.8),
    log(max(dat$Population) * 1.5),
    0
  ),

  fn = negLL,
  dat = dat,

  method = "BFGS",
  hessian = TRUE

)

fit$convergence

r_hat <- exp(fit$par[1])
K_hat <- exp(fit$par[2])
beta_hat <- fit$par[3]

parameter_estimates <- data.frame(
  Parameter = c("r", "K", "beta"),
  Estimate = c(r_hat, K_hat, beta_hat)
)

kable(parameter_estimates, digits = 3)

```

# Reconstruct Biomass Trajectory

```{r}

PredictedBiomass <- numeric(nrow(dat))

PredictedBiomass[1] <- dat$Population[1]

for(i in 1:(nrow(dat) - 1)) {

  Fnow <-
    q_hat *
    exp(-d_hat * dat$DaysFished[i]) *
    dat$DaysFished[i]

  r_eff <-
    r_hat *
    exp(beta_hat * dat$TempIndex[i])

  PredictedBiomass[i + 1] <-
    PredictedBiomass[i] +
    r_eff *
    PredictedBiomass[i] *
    (1 - PredictedBiomass[i] / K_hat) -
    Fnow *
    PredictedBiomass[i]

}

dat$PredictedBiomass <- PredictedBiomass

```

# Biomass Fit

```{r}

ggplot(dat, aes(Year)) +

  geom_line(
    aes(
      y = Population,
      colour = "Observed"
    ),
    linewidth = 1.2
  ) +

  geom_line(
    aes(
      y = PredictedBiomass,
      colour = "Predicted"
    ),
    linewidth = 1.2
  ) +

  scale_colour_manual(
    values = c(
      Observed = "black",
      Predicted = "blue"
    )
  ) +

  labs(
    title = "Observed vs Predicted Biomass",
    y = "Biomass",
    colour = ""
  )

```

# Precautionary Approach Reference Points

For the Schaefer model:

\[
B_{MSY} = \frac{K}{2}
\]

\[
LRP = 0.4 B_{MSY}
\]

\[
USR = 0.8 B_{MSY}
\]

```{r}

Bmsy <- K_hat / 2

LRP <- 0.40 * Bmsy

USR <- 0.80 * Bmsy

reference_points <- data.frame(
  Reference = c(
    "LRP",
    "USR",
    "BMSY"
  ),
  Value = c(
    LRP,
    USR,
    Bmsy
  )
)

kable(reference_points, digits = 0)

```

# Removal Reference

For a Schaefer model:

\[
F_{MSY} = \frac{r}{2}
\]

Removal Reference:

\[
RR = F_{MSY}
\]

or

\[
RR = \frac{r}{4}
\]

```{r}

Fmsy <- r_hat / 2

RR <- Fmsy 

RR

```

# Fishing Mortality Relative to RR

```{r}

dat <- dat %>%
  mutate(
    ExceedsRR = F_est > RR
  )

mean(dat$ExceedsRR)

ggplot(dat, aes(Year, F_est)) +

  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +

  geom_hline(
    yintercept = RR,
    colour = "red",
    linewidth = 1,
    linetype = 2
  ) +

  labs(
    title = "Fishing Mortality Relative to Removal Reference",
    y = "Fishing Mortality (F)"
  )

```

# Current Stock Status

```{r}

currentB <- tail(dat$Population, 1)

status <- case_when(
  currentB < LRP ~ "Critical",
  currentB < USR ~ "Cautious",
  TRUE ~ "Healthy"
)

status

```

# Biomass Status Plot

```{r}

ggplot(
  dat,
  aes(Year, Population)
) +

  annotate(
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = 0,
    ymax = LRP,
    fill = "red",
    alpha = 0.2
  ) +

  annotate(
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = LRP,
    ymax = USR,
    fill = "orange",
    alpha = 0.2
  ) +

  annotate(
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = USR,
    ymax = Inf,
    fill = "green",
    alpha = 0.15
  ) +

  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +

  labs(
    title = "Biomass Relative to PA Reference Points",
    y = "Biomass"
  )

```

# Stochastic Projection

```{r}

nsim <- 5000
nyears <- 20

futureB <- matrix(
  NA,
  nrow = nsim,
  ncol = nyears
)

futureB[,1] <- currentB

meanTemp <- mean(dat$TempIndex)
sdTemp <- sd(dat$TempIndex)

meanEffort <- mean(
  tail(dat$DaysFished, 5)
)

for(s in 1:nsim){

  for(y in 2:nyears){

    Tnow <- rnorm(
      1,
      meanTemp,
      sdTemp
    )

    r_eff <-
      r_hat *
      exp(beta_hat * Tnow)

    Bprev <- futureB[s, y - 1]

    Fproj <-
      q_hat *
      exp(-d_hat * meanEffort) *
      meanEffort

    Bnew <-
      Bprev +
      r_eff *
      Bprev *
      (1 - Bprev / K_hat) -
      Fproj *
      Bprev

    Bnew <- max(Bnew, 1)

    futureB[s, y] <-
      Bnew *
      rlnorm(
        1,
        meanlog = 0,
        sdlog = 0.15
      )
  }
}

```

# PA Risk Metrics

```{r}

risk_table <- data.frame(

  Year = 1:nyears,

  ProbAboveLRP =
    colMeans(futureB > LRP),

  ProbAboveUSR =
    colMeans(futureB > USR),

  ProbBelowLRP =
    colMeans(futureB < LRP)

)

kable(risk_table, digits = 3)

```
# Economic Viability Assumptions
```{r}
fleet_size <- 100

# Minimum average landings required
# for a vessel to remain viable

viability_threshold <- 2
```

# Projection Plot

```{r}

proj <- as.data.frame(futureB)

proj$Sim <- 1:nrow(proj)

proj_long <-
  pivot_longer(
    proj,
    cols = -Sim,
    names_to = "Year",
    values_to = "Biomass"
  )

proj_long$Year <-
  as.numeric(
    gsub("V", "", proj_long$Year)
  )

ggplot(
  proj_long,
  aes(
    Year,
    Biomass,
    group = Sim
  )
) +

  geom_line(alpha = 0.01) +

  geom_hline(
    yintercept = LRP,
    colour = "red"
  ) +

  geom_hline(
    yintercept = USR,
    colour = "orange"
  ) +

  labs(
    title = "20-Year Stochastic Biomass Projections",
    y = "Biomass"
  )

```

# Viability Summary

```{r}

terminalB <- futureB[, nyears]

viability <- data.frame(

  Metric = c(
    "Mean Biomass",
    "Median Biomass",
    "P(B > LRP)",
    "P(B > USR)",
    "P(B < LRP)"
  ),

  Value = c(
    mean(terminalB),
    median(terminalB),
    mean(terminalB > LRP),
    mean(terminalB > USR),
    mean(terminalB < LRP)
  )

)

kable(viability, digits = 3)

```

# Management Advice

```{r}

advice <- case_when(

  status == "Healthy" ~
    "Stock currently resides in the Healthy Zone.",

  status == "Cautious" ~
    "Stock currently resides in the Cautious Zone. Rebuilding actions should be considered.",

  TRUE ~
    "Stock currently resides in the Critical Zone. Rebuilding measures are recommended."

)

advice

```

# Assessment Summary

This assessment provides:

1. Seasonal temperature effects.
2. Effort-dependent depletion effects.
3. Biomass reference points (LRP and USR).
4. Removal reference based on FMSY/2.
5. Stock status relative to the PA framework.
6. Fishing mortality status relative to RR.
7. Twenty-year stochastic projections.
8. Probability of remaining above LRP and USR.