---
title: "Viability Kernel Analysis for Lobster Fishery Objectives"
author: ''
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

This document implements a **viability kernel (VK) analysis** for a lobster
fishery, exploring feasibility across multiple objectives:

- Fishing effort (days fished)
- Catch per trip (temperature dependent)
- Fuel and bait costs
- Economic performance (profit target)
- Environmental variability (temperature)

The viability kernel identifies combinations of **effort and temperature** for
which *all objectives can be met simultaneously*.

---

# Load packages

```{r}
library(dplyr)
library(ggplot2)
```

---

# Temperature-dependent catch per trip

```{r}
cpue_temp <- function(T,
                      T_opt = 9,
                      max_cpue = 120,
                      width = 2) {
  max_cpue * exp(-((T - T_opt)^2) / (2 * width^2))
}
```

---

# Define the state space

```{r}
vk_states <- expand.grid(
  days_fished = seq(50, 200, by = 5),
  temp = seq(6, 12, by = 0.25)
)
```

---

# Define economic parameters and targets

```{r}
price_per_lb <- 10
fuel_burn    <- 120
fuel_price   <- 1.80
bait_use     <- 40
bait_price   <- 1.25

cpue_target   <- 90
profit_target <- 2000
max_days      <- 180
```

---

# Compute derived metrics

```{r}
vk_states <- vk_states %>%
  mutate(
    cpue = cpue_temp(temp),
    revenue_trip = cpue * price_per_lb,
    cost_trip = fuel_burn * fuel_price + bait_use * bait_price,
    profit_trip = revenue_trip - cost_trip,
    seasonal_profit = profit_trip * days_fished
  )
```

---

# Define viability constraints

```{r}
vk_states <- vk_states %>%
  mutate(
    effort_ok = days_fished <= max_days,
    cpue_ok   = cpue >= cpue_target,
    profit_ok = seasonal_profit >= profit_target,
    viable    = effort_ok & cpue_ok & profit_ok
  )
```

---

# Viability kernel (Effort × Temperature)

```{r}
ggplot(vk_states, aes(days_fished, temp, fill = viable)) +
  geom_tile() +
  scale_fill_manual(values = c("grey80", "darkgreen")) +
  labs(
    title = "Viability Kernel for Lobster Fishery",
    x = "Days fished per season",
    y = "Temperature (°C)",
    fill = "State"
  ) +
  theme_minimal()
```

---

# Binding constraint diagnostics

```{r}
vk_states <- vk_states %>%
  mutate(
    failure_reason = case_when(
      !cpue_ok   ~ "Catch target not met",
      !profit_ok ~ "Profit constraint",
      !effort_ok ~ "Effort limit exceeded",
      TRUE       ~ "Viable"
    )
  )
```

```{r}
ggplot(vk_states, aes(days_fished, temp, fill = failure_reason)) +
  geom_tile() +
  labs(
    title = "Binding Constraints Across State Space",
    x = "Days fished per season",
    y = "Temperature (°C)",
    fill = "Constraint"
  ) +
  theme_minimal()
```

---

# Trade-offs: Temperature vs Profit

```{r}
ggplot(vk_states, aes(temp, seasonal_profit, color = viable)) +
  geom_point(alpha = 0.4) +
  scale_color_manual(values = c("grey60", "darkgreen")) +
  labs(
    title = "Seasonal Profit vs Temperature",
    x = "Temperature (°C)",
    y = "Seasonal profit ($)",
    color = "Viable"
  ) +
  theme_minimal()
```

---

# Extract the Viability Kernel

```{r}
vk_kernel <- filter(vk_states, viable)

range(vk_kernel$days_fished)
range(vk_kernel$temp)
```

---
