---
title: "Viability Kernel Analysis for Lobster Fishery Objectives"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
  toc_depth: 2
  number_sections: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
echo = TRUE, 
warning = FALSE, 
message = FALSE
)
```

# Introduction

This document implements a viability kernel (VK) analysis for a lobster fishery
under multiple objectives involving effort, economics, and climate variability.

# Load packages

```{r}
library(dplyr)
library(ggplot2)
```

# Temperature–dependent catch per trip

```{r}
cpue_temp <- function(T, T_opt = 9, max_cpue = 120, width = 2) {
  max_cpue * exp(-((T - T_opt)^2) / (2 * width^2))
}
```

# Define the state space

```{r}
vk_states <- expand.grid(
  days_fished = seq(50, 200, by = 5),
  temp = seq(6, 12, by = 0.25)
)
```

# Economic parameters and targets

```{r}
price_per_lb <- 10
fuel_burn <- 120
fuel_price <- 1.80
bait_use <- 40
bait_price <- 1.25

cpue_target <- 90
profit_target <- 2000
max_days <- 180
```

# Derived metrics

```{r}
vk_states <- vk_states %>%
  mutate(
    cpue = cpue_temp(temp),
    revenue_trip = cpue * price_per_lb,
    cost_trip = fuel_burn * fuel_price + bait_use * bait_price,
    profit_trip = revenue_trip - cost_trip,
    seasonal_profit = profit_trip * days_fished
  )
```

# Viability constraints

```{r}
vk_states <- vk_states %>%
  mutate(
    effort_ok = days_fished <= max_days,
    cpue_ok = cpue >= cpue_target,
    profit_ok = seasonal_profit >= profit_target,
    viable = effort_ok & cpue_ok & profit_ok
  )
```

# Viability Kernel (Effort × Temperature)

```{r}
ggplot(vk_states, aes(days_fished, temp, fill = viable)) +
  geom_tile() +
  scale_fill_manual(values = c("grey80", "darkgreen")) +
  theme_minimal()
```

# Seasonal extensions to the viability kernel

```{r}
seasons <- tibble(
  season = c("Early", "Mid", "Late"),
  price_mult = c(0.85, 1.00, 1.15),
  quality_mult = c(0.8, 1.0, 1.1),
  temp_shift = c(-0.5, 0, 0.5)
)
```

```{r}
cpue_temp_season <- function(T, season, T_opt = 9, max_cpue = 120, width = 2) {
  penalty <- case_when(
    season == "Early" ~ 0.8,
    season == "Mid" ~ 1.0,
    TRUE ~ 0.9
  )
  penalty * max_cpue * exp(-((T - T_opt)^2) / (2 * width^2))
}
```

```{r}
vk_seasonal <- expand.grid(
  days_fished = seq(50, 200, 5),
  temp = seq(6, 12, 0.25),
  season = seasons$season
) %>% left_join(seasons, by = "season")
```

```{r}
vk_seasonal <- vk_seasonal %>%
  mutate(
    cpue = cpue_temp_season(temp + temp_shift, season),
    effective_price = price_per_lb * price_mult * quality_mult,
    revenue_trip = cpue * effective_price,
    cost_trip = fuel_burn * fuel_price + bait_use * bait_price,
    profit_trip = revenue_trip - cost_trip,
    seasonal_profit = profit_trip * days_fished,
    viable_season = cpue >= cpue_target & seasonal_profit >= profit_target & days_fished <= max_days
  )
```

```{r}
vk_kernel_strict <- vk_seasonal %>%
  group_by(days_fished, temp) %>%
  summarise(viable = all(viable_season), .groups = "drop")
```

```{r}
ggplot(vk_kernel_strict, aes(days_fished, temp, fill = viable)) +
  geom_tile() +
  scale_fill_manual(values = c("grey80", "darkgreen")) +
  theme_minimal()
```
