
```{r,include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, comment = FALSE, error = FALSE}
setwd("C:/Users/HowseVJ/Documents/GitHub/bio.lobster/inst/Updates/LFA35-38/LFA35_Update")

require(rgdal)
require(devtools)
require(roxygen2)
require(geosphere)
require(SpatialHub)
require(lubridate)
require(bio.utilities)
require(bio.lobster)
require(rstanarm)
require(rlang)
require(glue)
require(PBSmapping)

 
	p = bio.lobster::load.environment()
	la()
	assessment.year = 2023 ##Check Year
	p$syr = 1989
  p$yrs = p$syr:assessment.year



figdir = file.path("LFA35_Update")
    
	    p$lfas = c("35") # specify lfa

	   

## from Oracle
#lobster.db('logs.redo') 
#lobster.db('temperature.data.redo')
lobster.db('temperature.data')
#groundfish.db('odbc.redo',datayrs = 1970:2022)
#logs=lobster.db('process.logs.redo')
logs=lobster.db("process.logs")
#land = lobster.db('annual.landings.redo')
land =lobster.db("annual.landings")
#Sland=lobster.db('seasonal.landings.redo')
Sland = lobster.db('seasonal.landings')


```
# Analysis and Response

## Indicators of Stock Status

The stock status of the Lobster in LFA 35 is assessed using primary, secondary, and contextual
indicators. This update includes the primary indicator that is used to define stock status in
relation to reference points defined in Cook et al. (In press.1) and secondary indicators that
display time-series trends but do not have reference points. The data sources available for
establishing indicators for LFA 35 come from both fishery-dependent and fishery-independent
data sources. Fishery-dependent data consist of commercial logbooks that report information on
date, location (grid), effort, and estimated catch. The fishery-independent data sources are from
the DFO Maritimes Region Summer Research Vessel Survey (herein RV survey), and the Scallop Survey. 
Indicators from surveys updated where possible.


# Primary Indicator

In LFA 35, there is one primary indicator for stock status that describes the time-series trends
relative to reference points. The primary indicator for describing stock status is standardized
commercial Catch Per Unit Effort (CPUE). There is currently no primary indicator of fishing
pressure or exploitation.


## Catch Per Unit Effort

Commercial catch rates are a preferred indicator over landings, as they are standardized to
account for the level of fishing effort. This is especially important in effort-controlled fisheries.
The commercial fishing data used in the estimation of catch rates were obtained from mandatory logbooks that were put in place in the mid-2000s.
It has been well documented that trap-based catch rates vary throughout a fishing season due to factors other than available
biomass. These factors include fishing behavior, localized depletion, and environmental
conditions (Drinkwater et al. 2006). In an effort to account for these factors, CPUE data were
standardized through generalized linear modelling with explanatory variables of Year, Day of
Season, Temperature, and the interaction between Day of Season and Temperature. Year was
treated as a factor rather than a continuous variable to reduce any constraints from inter-annual
variability.

Model predictions were made for day 1 of the fishing season at the median day 1 temperature
across all years. The available time series covers the current high-productivity period and a
lower-productivity period from 2006–2010 (Figure \@ref(fig:CPUE35)). The median of the high-productivity period
(2011–2018) was used as the proxy for the biomass at carrying capacity (K). Following the
recommendations of DFO (2009), the Upper Stock Reference (USR) and Limit Reference Point
(LRP) were set to 40% and 20% of the K proxy. A 3-year running median is used to compare
the standardized CPUE to the USR and LRP. This value will dampen the impact of any
anomalous years that may occur due to factors outside of changes in abundance.

```{r CPUE35, echo=FALSE, results=FALSE, fig.cap="Time series of standardized commercial catch rates (kg/trap haul; black dots) for LFA 35, along with the 3-year running median (solid blue line). The horizontal lines represent the Upper Stock Reference (dashed green line) and Limit Reference Point (dotted red line). The data for 2021-2022 fishing season are incomplete (blue triangle)."}

	logs=lobster.db("process.logs")
	TempModelling = TempModel( annual.by.area=F, redo.data=F)
			CPUE.data<-CPUEModelData(p,redo=T,TempModelling)

			## Commercial CPUE MOdels
			mf1 = formula(logWEIGHT ~ fYEAR + DOS + TEMP + DOS * TEMP)

			CPUE.data<- CPUEModelData(p,redo=F)
			t=with(subset(CPUE.data,DOS==1),tapply(TEMP,LFA,mean))
		
			pData=list()

			CPUEModelResults = list()

			for(i in 1:length( p$lfas)){

				mdata = subset(CPUE.data,LFA==p$lfas[i]&SYEAR%in%p$yrs)
				if(nrow(mdata)>10){
				CPUEModelResults[[i]] = CPUEmodel(mf1,mdata,t=t[i],d=1)
				pData[[i]]=CPUEModelResults[[i]]$pData
				pData[[i]]$LFA=p$lfas[i]
			}
			}

			names(CPUEModelResults) = p$lfas

			CPUEindex=do.call("rbind",pData)


	# plot

	#tiff("C:/Users/Howsevj/Documents/3638_Update/36plotcpue.tiff", width=7,height=6, units="in",res=300)

	tiff("R:/Shared/CSA_Office/2022 CSAS Processes/LFA 35 Update/Completed Edits and FR Fig/36plotcpue.tiff", width=7,height=6, units="in",res=300)
	l35 = subset(CPUEindex,LFA==35,c("YEAR","mu"))
	k = median(l35$mu[which(l35$YEAR %in% 2011:2018)])
	usr = .4*k
	lrp = .2*k
	

plot(l35$YEAR,l35$mu,xlab='Année',ylab='CPUE Standardisé (kg/casier levé)',type='p',pch=16,ylim=c(0,6),xlim=c(2005,2022))
	running.median = with(rmed(l35$YEAR,l35$mu),data.frame(YEAR=yr,running.median=x))
	l35=merge(l35,running.median,all=T)
	lines(l35$YEAR,l35$running.median,col='blue',lty=1,lwd=3)
	points(l35$YEAR[nrow(l35)],l35$mu[nrow(l35)], type='b', pch=24,bg='blue')
	abline(h=usr,col='green',lwd=2,lty=2)
	abline(h=lrp,col='red',lwd=2,lty=3)
	
dev.off()
```


```{r RAWCPUE35, echo=FALSE, results=FALSE, fig.cap="Time series of raw commercial catch rates (kg/trap haul; black dots) for LFA 35. The data for 2021-2022 fishing season are incomplete (blue triangle)."}

### USING RAW UNPROCESSED LOGS

a = lobster.db('process.logs')
a = subset(a,LFA==35)
b=aggregate(cbind(NUM_OF_TRAPS,WEIGHT_KG)~SYEAR,data=a,FUN=sum)
b$CPUE=b$WEIGHT_KG/b$NUM_OF_TRAPS

tiff("R:/Shared/CSA_Office/2022 CSAS Processes/LFA 35 Update/Completed Edits and FR Fig/36plotcpueraw.tiff", width=7,height=6, units="in",res=300)
## Raw CPUE
plot(b$SYEAR, b$CPUE, xlab="Année", ylab="CPUE (kg/casier levé)", type='p', pch=16, ylim=c(0,2.5))
	points(b$SYEAR[nrow(b)],b$CPUE[nrow(b)], type='b', pch=24,cex = 1.5, bg='blue')
			running.median = with(rmed(b$SYEAR,b$CPUE),data.frame(SYEAR=yr,running.median=x))
	b=merge(b,running.median,all=T)
	lines(b$SYEAR,b$running.median,col='blue',lty=1,lwd=3)
dev.off()
```


The trend in the standardized CPUE indicates that an increase in stock biomass occurred between 2005 and
2012 ( Figure \@ref(fig:CPUE35) ). Standardized CPUE has remained high (more than twice the USR) since 2011. The
3-year running median for CPUE for the 2022 season is 4.72 kg per Trap Haul (kg/TH), which is
above the USR (1.94 kg/TH) and LRP (0.97 kg/TH). The Standardized CPUE value for 2022 is preliminary due to outstanding logs; as of August 23rd the monthly reporting rate was between 44% to 88% by month. Temperature data incorporated into the standardized CPUE estimates was limited for 2022 at the time of the update so the model CPUE model only includes 2021 temperature data. Raw CPUE, calculated from unprocessed logs was included to highlight the variability in fishing activity and catch over the time series ( Figure \@ref(fig:RAWCPUE35) ).



# Secondary Indicators

Secondary indicators represent time-series trends that are tracked individually, without defined
reference points. The secondary indicators for LFAs 35 include the LFA-specific
landings and total effort, recruit abundance, commercial biomass, and relative fishing mortality
estimates from the DFO RV survey Bay of Fundy region (strata 484, 490–495). Additionally, recruit
abundance from the Scallop survey will be included in this update for the first time since the assessment (DFO 2021). 

## Landings and Effort

Commercial landings are related to population biomass, as fishery controls are input- (effort
controls) rather than output-based (total allowable catch). There are many factors that can affect
this relationship, including changes in fishing effort, catchability (including the effects of
environment, gear efficiency), Lobster size distribution, and the spatial overlap between
distribution of Lobster and effort.

Fishing effort, recorded as the number of Trap Hauls (THs), in the Lobster fishery is controlled
by fishing season length, trap limits, and limited number of fishing licences. Consequently, there
is a maximum fishing effort that can be deployed; however, this maximum is never met because
factors such as weather conditions, seasonally variable catch rates, and fishing partnerships
limit the total number of THs. Total fishing effort is calculated from mandatory logbooks.

```{r landeffort,echo=FALSE, results=FALSE, fig.cap="Time series of landings (bars), and effort (sold line with points) for LFA 35. The data for 2021-2022 fishing season is incomplete (blue bar for landings, black triangle for effort))."}

a=lobster.db('process.logs')
a = subset(a,LFA==35)
b=aggregate(cbind(NUM_OF_TRAPS,WEIGHT_KG)~SYEAR,data=a,FUN=sum)
b$CPUE=b$WEIGHT_KG/b$NUM_OF_TRAPS
h=lobster.db('seasonal.landings.redo')
h=lobster.db('seasonal.landings')
h
h=h[,c('SYEAR','LFA35')]
h
h$SYEAR=1976:2023
merge(b,h)
g=merge(b,h)
g$EFF = g$LFA35/g$CPUE

tiff("R:/Shared/CSA_Office/2022 CSAS Processes/LFA 35 Update/Completed Edits and FR Fig/36plotlandeff.tiff", width=7,height=6, units="in",res=300)
par(mar=c(5.1, 4.1, 4.1, 5.1),las=1)

plot(g$SYEAR,g$LFA35,xlab='Année',ylab='Débarquements (t)',type='h',ylim=c(0,max(g$LFA35)*1.2),pch=15,col='grey',lwd=10,lend=3)
lines(g$SYEAR[nrow(g)],g$LFA35[nrow(g)],type='h',pch=21,col='steelblue4',lwd=10,lend=3)
par(new=T)

plot(g$SYEAR,g$EFF,ylab='',xlab='', type='b', pch=16, axes=F,ylim=c(0,max(g$EFF,na.rm=T)))
points(g$SYEAR[nrow(g)],g$EFF[nrow(g)], type='b', pch=24,bg='black')
axis(4)
mtext("Effort (milliers de casiers levés)", 4, 3.5, outer = F,las=0)	

dev.off()		


subsamp<-g[g$SYEAR >2009 & g$SYEAR <2023,]
range(subsamp$LFA35)
median(subsamp$LFA35)

```

Historical landings in LFA 35 between 1947 and 1984 had a median of 134 t with a range of
75–184 t, increasing slightly between 1984 and 1994 to a median of 250.5 t (range 226–330 t).
In the more recent years, LFA 35 landings have increased substantially with a median from
2010–2022 of 3,072 t (range 1,829 to 3,941 t). Landings have decreased in recent years from the 3631 t reported in 2018.
Landings for the 2021–2022 season are currently 1829 t but do not represent the total for the season due to the outstanding logs previously
mentioned ( Figure \@ref(fig:landeffort) ). 

## Scallop Survey Recruit Abundance

Annual surveys for Sea Scallops have been conducted since the early 1980s to assess
abundance (Sameoto et al. 2012, Smith et al. 2012). These surveys started in the Bay of Fundy
in 1981 and were extended into the area off southwest Nova Scotia in 1991. Lobster are caught as a bycatch in a subset of
the stations and are measured prior to being returned to the ocean.Scallops are typically found on gravel sea bottoms, a habitat not favored by Lobster (Tremblay et al. 2009), but the two species do overlap in some areas. The surveys are primarily conducted in July for the Bay of Fundy (LFAs 35 and 36) and August around Grand Manan Island (LFA
38), which, in some years, overlaps the Lobster moulting period. Scallop dredges tend to
capture Lobster that are under the legal size and are useful as indicators of recruitment.

The abundance of Lobster recruits (70–82 mm carapace length) from DFO Scallop Survey tows within LFA 35 had been
increasing after 2008, however in the most recent years have shown a decline (2019 and 2021), the scallop survey was not completed in 2020 due to the COVID-19
global pandemic (Figure \@ref(fig:Scalrecruit)) , and data from the 2022 survey was not yet available at the time of this update. 

```{r Scalrecruit,echo=FALSE, results=FALSE, fig.cap="Time series of recruit abundance from the DFO Scallop Survey in LFA 35.The red line represents the 3-year running median."}
p = bio.lobster::load.environment()
la()

assessment.year = 2021 ########### check the year ############### !!!!!!!!!!!

p$syr = 1989
p$yrs = p$syr:assessment.year



# define place for figures to go
#figdir = file.path("C:/Users/HowseVJ/Documents/GitHub/bio.lobster/inst/Updates/LFA35-38")

p$lfas = c("35") # specify lfa
p$subareas = c("35") # specify subareas for data summary

# update data through RODBC

#lobster.db('scallop.redo')
lobster.db('scallop')

scalSurv<-ScallopSurveyProcess(size.range=c(70,82.5),bin.size=2.5)
scalSurv<-scalSurv[scalSurv$YEAR <2022,] # 2022 data not complete yet

R1.ab.35=with(subset(scalSurv,LFA==35&YEAR>1998),tapply(LobDen,YEAR,mean,na.rm=T))
R1.ab.35_19=with(subset(scalSurv,LFA==35&YEAR>1998&YEAR<2020),tapply(LobDen,YEAR,mean,na.rm=T))



require(dplyr)
RA35<-as.data.frame(R1.ab.35)
RA35<-data.frame(cbind(rownames(RA35), data.frame(RA35, row.names = NULL)))
RA35<-RA35 %>%
  rename(Year = rownames.RA35., 
          Recruit_Abundance = R1.ab.35)

tiff("R:/Shared/CSA_Office/2022 CSAS Processes/LFA 35 Update/Completed Edits and FR Fig/35plotrecruit.tiff", width=7,height=6, units="in",res=300)
plot(RA35$Year, RA35$Recruit_Abundance, pch=16,xlab='Année',ylab = 'Abondance des recrues')
lines(1999:2019,runmed(R1.ab.35_19,3),lwd=2, col='salmon')
dev.off()

```






```{r Stratified Analysis Commercial, eval=FALSE, echo=FALSE, results=FALSE,fig.cap="Time Series of RV survey trends for LFA 35-38 commercial biomass. Values prior to 1999 (open circles) were derived using the mean proportion of commercial to total biomass between 1999 and 2018 (0.746), error bars are 95% bootstrapping confidence intervals. The red line represents the 3-year running median" }
require(bio.survey)
require(bio.lobster)

p = bio.lobster::load.environment()
p$libs = NULL
ff = "LFA35-38Assessment"
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
dir.create(fpf1,showWarnings=F)
dir.create(fp1,showWarnings=F)

p$yrs = 1970:2022
p1 = p


stratifiedAnalysesCommercial = function( p=p1, survey,lfa, fpf = fpf1, fp = fp1,f=ff,wd=10,ht=8){
    p$series =c('summer')# p$series =c('georges');p$series =c('fall')
    p$years.to.estimate = p$yrs
    p$length.based = T
    p$by.sex = T
    p$size.class = c(83,300)
    p$sex = c(1,2)
    p$bootstrapped.ci=T
    p$strata.files.return=F
    p$vessel.correction.fixed=1.2
    p$strat = NULL
    p$clusters = c( rep( "localhost", 7) )
    p$strata.efficiencies = F
    p = make.list(list(yrs=p$years.to.estimate),Y=p)
    p$define.by.polygons = T
    p$lobster.subunits=F
    p$area = lfa
    p$reweight.strata = T #this subsets 

    aout= dfo.rv.analysis(DS='stratified.estimates.redo',p=p)
    write.csv(aout,file=file.path(fpf, paste(lfa,'DFOCommB.csv')))

      p$add.reference.lines = F
    p$time.series.start.year = p$years.to.estimate[1]
    p$time.series.end.year = p$years.to.estimate[length(p$years.to.estimate)]
    p$metric = 'weights' #weights
    p$measure = 'stratified.total' #'stratified.total'
    p$figure.title = ""
    p$reference.measure = 'median' # mean, geomean
    p$file.name =  file.path(f,paste(lfa,'DFOrestratifiedtotalweightscommercial.png',sep=""))

    p$y.maximum = NULL # NULL # if ymax is too high for one year
    p$show.truncated.weights = F #if using ymax and want to show the weights that are cut off as values on figure

    p$legend = FALSE
    p$running.median = T
    p$running.length = 3
    p$running.mean = F #can only have rmedian or rmean
     p$error.polygon=F
    p$error.bars=T
    require(bio.lobster)
     p$ylim2 = NULL
    xx = aggregate(ObsLobs~yr,data=aout,FUN=sum)
    names(xx) =c('x','y')
     p$ylim=NULL
     ref.out= figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
     return(aout)
  }



aout = stratifiedAnalysesCommercial(survey='DFO',lfa='LFA35-38')
write.csv(aout,file.path(fp1,'LFA3538CommercialB.csv'))
```


```{r Stratified Analysis,  eval=FALSE, echo=FALSE, results=FALSE,fig.cap="  "}
require(bio.survey)
require(bio.lobster)
require(lubridate)
require(devtools)
require(bio.utilities)
require(PBSmapping)
la()
p = bio.lobster::load.environment()
p$libs = NULL
ff = "LFA35-38Assessment"
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
dir.create(fpf1,showWarnings=F)
dir.create(fp1,showWarnings=F)
p1 = p
p1$yrs = 1970:2022

stratifiedAnalyses = function(p=p1, survey,lfa, fpf = fpf1, fp = fp1,f=ff,wd=10,ht=8){
                p$series =c('summer')
                p$define.by.polygons = T
                p$lobster.subunits=F
                p$area = lfa
                p$years.to.estimate = p$yrs[-1]
                p$length.based = F
                p$by.sex = F
                p$bootstrapped.ci=T
                p$strata.files.return=F
                p$vessel.correction.fixed=1.2
                p$strat = NULL
                p$clusters = c( rep( "localhost", 7) )
                p$strata.efficiencies = F
                p = make.list(list(yrs=p$years.to.estimate),Y=p)
                p$reweight.strata = T #this subsets 
            

      aout= dfo.rv.analysis(DS='stratified.estimates.redo',p=p)
      write.csv(aout,file=file.path(fpf, paste(lfa,'DFOtotalabund.csv')))


                              p$add.reference.lines = F
                              p$time.series.start.year = p$years.to.estimate[1]
                              p$time.series.end.year = p$years.to.estimate[length(p$years.to.estimate)]
                              p$metric = 'numbers' #weights
                              p$measure = 'stratified.mean' #'stratified.total'
                              p$figure.title = ""
                              p$reference.measure = 'median' # mean, geomean
                              p$file.name = file.path(f,paste(lfa,'DFOrestratifiednumbers.png',sep=""))

                          p$y.maximum = NULL # NULL # if ymax is too high for one year
                        p$show.truncated.numbers = F #if using ymax and want to show the numbers that are cut off as values on figure

                                p$legend = FALSE
                                p$running.median = T
                                p$running.length = 3
                                p$running.mean = F #can only have rmedian or rmean
                               p$error.polygon=F
                              p$error.bars=T

                               p$ylim=NULL
                      if(lfa == 'LFA35-38') p$ylim=c(0,150)
                       p$box=T
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiednumbersNOY.png', sep=""))
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)

                       p$box=T
                        p$ylim=NULL
                       p$metric = 'weights'
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiedweightsNOY.png',sep=""))
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)

                       p$box=NULL
                       p$ylim=NULL
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiedDWAO.png',sep=""))
                       p$metric = 'dwao'
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
                       
                       p$file.name = file.path(f,paste(lfa,'DFOrestratifiedgini.png',sep=""))
                       p$metric = 'gini'
                       p$ylim =c(0,1)
                       ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
                       p$ylim = NULL
                       return(aout)
        }

    
a = stratifiedAnalyses(survey='DFO',lfa='LFA35-38')

```


```{r Stratified Analysis Recruits, eval=FALSE, echo=FALSE, results=FALSE,fig.cap="  "}
require(bio.survey)
require(bio.lobster)
require(lubridate)
require(devtools)
require(bio.utilities)
require(PBSmapping)
la()
p = bio.lobster::load.environment()
p$libs = NULL
ff = "LFA35-38Assessment"
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
dir.create(fpf1,showWarnings=F)
dir.create(fp1,showWarnings=F)
p1 = p
p1$yrs = 1970:2022

stratifiedAnalysesRecruits = function(p=p1, survey,lfa, fpf = fpf1, fp = fp1,f=ff,wd=10,ht=8){
      p$series =c('summer')# p$series =c('georges');p$series =c('fall')
      p$area = lfa
      p$years.to.estimate = p$yrs
      p$length.based = T
      p$by.sex = T
      p$size.class = c(70,82)
      p$sex = c(1,2)
      p$bootstrapped.ci=T
      p$strata.files.return=F
      p$vessel.correction.fixed=1.2
      p$strat = NULL
      p$clusters = c( rep( "localhost", 7) )
      p$strata.efficiencies = F
      p = make.list(list(yrs=p$years.to.estimate),Y=p)
      p$define.by.polygons = T
      p$lobster.subunits=F
      p$reweight.strata = T #this subsets 
  
    aout= dfo.rv.analysis(DS='stratified.estimates.redo',p=p)
     write.csv(aout,file=file.path(fpf, paste(lfa,'DFOrecruits.csv')))

                        p$add.reference.lines = F
                              p$time.series.start.year = p$years.to.estimate[1]
                              p$time.series.end.year = p$years.to.estimate[length(p$years.to.estimate)]
                              p$metric = 'numbers' #numbers
                              p$measure = 'stratified.mean' #'stratified.total'
                              p$figure.title = ""
                              p$reference.measure = 'median' # mean, geomean
                              p$file.name =  file.path(f,paste(lfa,'DFOrestratifiednumbersrecruits.png',sep=""))
                          p$y.maximum = NULL # NULL # if ymax is too high for one year
                        p$show.truncated.numbers = F #if using ymax and want to show the numbers that are cut off as values on figure
                                p$legend = FALSE
                                p$running.median = T
                                p$running.length = 3
                                p$running.mean = F #can only have rmedian or rmean
                               p$error.polygon=F
                              p$error.bars=T
                      if(lfa == 'LFA35-38') p$ylim = c(0,80)
                      p$file.name =  file.path(f,paste(lfa,'NOYDFOrestratifiednumbersrecruits.png',sep=""))
                      ref.out=   figure.stratified.analysis(x=aout,out.dir = 'bio.lobster', p=p,wd=wd,ht=ht)
            }

stratifiedAnalysesRecruits(survey='DFO',lfa='LFA35-38')



```


```{r Landings Survey Relative F,  eval=FALSE, echo=FALSE, results=FALSE,fig.cap="  "}

require(bio.lobster)
require(PBSmapping)

a = lobster.db('annual.landings')
b = lobster.db('seasonal.landings')
ff = "LFA35-38Assessment"
 
b$YR = substr(b$SYEAR,6,9)
a = subset(a,YR<1976)
b = subset(b,YR>1975 & YR<=2022)
fpf1 = file.path(project.figuredirectory('bio.lobster'),ff)
fp1 = file.path(project.datadirectory('bio.lobster'),"analysis",ff)

LFA = c('LFA35','LFA36','LFA38')
for(i in LFA){
		aa = a[,c('YR',i)]
		bb = b[,c('YR',i)]
		aa = (rbind(aa,bb))
		aa = aa[order(aa$YR),]
		file.name = paste('Landings',i,'.png',sep="")
		png(file=file.path(fpf1,'LandingsL3538.png'),units='in',width=15,height=12,pointsize=18, res=300,type='cairo')
		plot(aa$YR,aa[,i],type='h',lwd=4,col='black',xlab='Year',ylab='Landings (t)')
		lines(aa$YR,runmed(aa[,i],3),col='salmon',lwd=3)
		dev.off()
		}


###lfa 35-38

df2 = read.csv(file.path(fpf1,'LFA35-38 DFOCommB.csv'))
df =  read.csv(file.path(fpf1,'LFA35-38 DFOtotalabund.csv'))
df = subset(df,yr<1999)
df2 = subset(df2,yr>1998)
df = as.data.frame(rbind(df,df2))
df = df[,c('yr','w.Yst','w.ci.Yst.l','w.ci.Yst.u')] #proportion of total weight that is commercial
df$w.Yst[which(df$yr<1999)] <- df$w.Yst[which(df$yr<1999)]*0.746
df$w.ci.Yst.l[which(df$yr<1999)] <- df$w.ci.Yst.l[which(df$yr<1999)]*0.746
df$w.ci.Yst.u[which(df$yr<1999)] <- df$w.ci.Yst.u[which(df$yr<1999)]*0.746



 	 #png(file=file.path(fpf1,'LFA35-38CommBDFOextended.png'),units='in',width=10,height=8,pointsize=18, res=300,type='cairo')
 	 with(df,plot(yr,w.Yst,pch=1,xlab='Year',ylab='Commerical Biomass (t)',ylim=c(0,9500)))
	 with(df,arrows(yr,y0=w.ci.Yst.u,y1=w.ci.Yst.l, length=0))
	 with(subset(df,yr>1998),points(yr,w.Yst,pch=16))
	 xx = rmed(df$yr,df$w.Yst)
	 xx = as.data.frame(do.call(cbind,xx))
	 with(subset(xx,yr<1999),lines(yr,x,col='salmon',lwd=1))
	 with(subset(xx,yr>1998),lines(yr,x,col='salmon',lwd=3))
	 dev.off()
	 
a$L3538 = rowSums(a[,12:14])
b$L3538 = rowSums(b[,4:6])
c358 = as.data.frame(rbind(a[,c('YR','L3538')],b[,c('YR','L3538')]))
names(c358)[1] = 'yr'
df  =merge(df,c358)
df$rL = df$L3538/(df$w.ci.Yst.l+df$L3538)
df$rU =df$L3538/ (df$w.ci.Yst.u+df$L3538)
df$rM = df$L3538/(df$w.Yst+df$L3538)


 	 png(file=file.path(fpf1,'LFA3538RelFDFO.png'),units='in',width=10,height=8,pointsize=18, res=300,type='cairo')
		plot(df$yr,df$rM,type='p',pch=16,col='black',xlab='Year',ylab='Relative F')
		arrows(df$yr,y0 = df$rL,y1 = df$rU,length=0)
		with(rmed(df$yr,df$rM),lines(yr,x,col='salmon',lwd=3))
dev.off()

```

